---
title: "CH variants analysis"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: no
    theme: cerulean
    highlight: pygments
    df_print: paged
editor_options:
  chunk_output_type: console
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 25px;
}

table {
    margin-top: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
```


<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">Potential CH variants analysis in GnomAD database</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(gtsummary)

theme_gtsummary_compact()
theme_set(theme_classic())
```

```{r load}
gnomad_variants <-
  read.delim(paste0(here::here(), "/chr2_CH_variants.vcf.gz"), 
             colClasses = c("ALT"="character"),
              sep = " ")
known_CH_genes <-
  read_table(paste0(here::here(), "/3390851_Covered.bed"), 
    col_names = FALSE, skip = 2)
known_CH_gene_list <- paste0(c(unique(known_CH_genes$X4)), collapse = "|")
```

We want to look at skewness of age distribution in gnomad. We don't have access to real numerical age in this data but was wondering what is the age distribution of Cosmic patients presenting variants found in gnomAD. Note : First plot - a patient can be represented multiple time (1 time for each variant the patient express). Second plot - patient represented once.  
Population age in chromosome `r as.character(gnomad_variants %>% distinct(X.CHROM))` (Cosmic samples)
```{r Population age}
gnomad_variants %>% 
  ggplot(aes(x= Age, fill= str_detect(SYMBOL, known_CH_gene_list)))+
  geom_histogram(binwidth = 1)+
  xlim(0, 100)+ 
  scale_fill_discrete(name = NULL, labels = c("Not CH genes", "CH genes"))

gnomad_variants %>% 
  distinct(Sample.name, .keep_all = TRUE) %>% 
  ggplot(aes(x= Age, fill= str_detect(SYMBOL, known_CH_gene_list)))+
  geom_histogram(binwidth = 1)+
  xlim(0, 100)+ 
  scale_fill_discrete(name = NULL, labels = c("Not CH genes", "CH genes"))
```


The number of variants in chromosome `r as.character(gnomad_variants %>% distinct(X.CHROM))` is `r nrow(gnomad_variants %>% distinct(IDs))`.
```{r BH}
pval_selection <- 0.05
mannw_variants <- gnomad_variants %>% 
  filter(adjusted_pval <= pval_selection)
```
In chromosome `r as.character(mannw_variants %>% distinct(X.CHROM))`, the number of variants with BH ≤ `r pval_selection` is `r nrow(mannw_variants %>% distinct(IDs))`.
```{r Hedges g}
hedges_selection <- "negligible|small"
hedges_variants <- mannw_variants %>%
  # filter(effect_size_cat != "negligible")
  filter(!str_detect(effect_size_cat, hedges_selection))
```
In chromosome `r as.character(hedges_variants %>% distinct(X.CHROM))`, the number of variants with Hedges g effect size better than `r hedges_selection %>% str_replace(., "\\|", " or ")` is `r nrow(hedges_variants %>% distinct(IDs))`.

```{r homozygous}
zygote_variants <- hedges_variants %>% 
  filter(nbr_hom_individuals == 0)
```
In chromosome `r as.character(zygote_variants %>% distinct(X.CHROM))`, the number of variants with NO homozygous individuals is `r nrow(zygote_variants %>% distinct(IDs))`.

```{r zygote}
CH_variants <- zygote_variants %>%
  filter(is_gencode_protein_coding_gene == "Yes")
```
In chromosome `r as.character(CH_variants %>% distinct(X.CHROM))`, the number of variants which are on a protein coding gene is `r nrow(CH_variants %>% distinct(IDs))`. These variants are located on `r nrow(CH_variants %>% distinct(IDs, .keep_all = TRUE) %>% distinct(SYMBOL))` genes.
<br>
<br>

# GnomAD Analysis for each variants
```{r}
CH_variants <- CH_variants %>% 
  mutate(known_CH_genes = case_when(
    str_detect(SYMBOL, known_CH_gene_list)          ~ "known CH gene",
    TRUE                                            ~ "not known CH gene"
  ))
```

## Number of variants per gene 
```{r fig number of variant per gene, fig.height=20, fig.width=12}
CH_variants %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
  ggplot(aes(x= SYMBOL, fill= str_detect(SYMBOL, known_CH_gene_list)))+
  geom_bar()+
  ggtitle("Number of variants per gene in chromosome 2")+ 
  scale_fill_discrete(name = NULL, labels = c("Not CH genes", "CH genes"))+
  labs(x= NULL, y="Number of variants")+
  coord_flip()
```


```{r fig number of variant per gene2, fig.height=15, fig.width=12}
CH_variants %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
  # mutate(known_CH_gene = fct_relevel(known_CH_gene, "known CH gene")) %>%
  ggplot(aes(x= SYMBOL, fill= str_detect(SYMBOL, known_CH_gene_list)))+
  geom_bar()+
  ggtitle("Number of variants per gene")+ 
  scale_fill_discrete(name = NULL, labels = c("Not CH genes", "CH genes"))+
  labs(x= NULL, y="Number of variants")+
  coord_flip()+
  facet_wrap(str_detect(SYMBOL, known_CH_gene_list) ~ ., 
             scales = "free")+ 
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )
```

## Number of overall variants per gene
Averaged Not a CH genes in 1 bar
```{r}
CH_variants %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
    
  mutate(known_CH_gene = case_when(
    str_detect(SYMBOL, known_CH_gene_list)          ~ SYMBOL,
    TRUE                                            ~ "not CH gene (average)"
  )) %>% select(IDs, SYMBOL, known_CH_gene) %>% 
  arrange(known_CH_gene) %>% 
  group_by(known_CH_gene) %>% 
  mutate(number_of_variants_per_gene = n()) %>% 
  distinct(SYMBOL, .keep_all = TRUE) %>% 
  select(-IDs) %>% 
  
  group_by(known_CH_gene) %>% 
  mutate(number_of_gene = n()) %>%
  ungroup() %>% 
  mutate(average_of_variants_per_gene = number_of_variants_per_gene / number_of_gene) %>%
  distinct(known_CH_gene, .keep_all = TRUE) %>% 

  mutate(known_CH_gene = fct_relevel(known_CH_gene, "not CH gene")) %>%
  ggplot(aes(x= known_CH_gene, y= average_of_variants_per_gene,
             fill= known_CH_gene == "not CH gene"
             ))+
  geom_bar(stat = "identity", show.legend = FALSE)+
  ggtitle("Number of variants per gene")+
  labs(x= NULL)+
  coord_flip()
```

Which gene shows high number of variants?
```{r}
CH_variants %>% 
  filter(!is.na(SYMBOL)) %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
    
  mutate(known_CH_gene = case_when(
    str_detect(SYMBOL, known_CH_gene_list)          ~ "known CH gene",
    TRUE                                            ~ SYMBOL
  )) %>% select(IDs, SYMBOL, known_CH_gene) %>% 
  arrange(known_CH_gene) %>% 
  group_by(known_CH_gene) %>% 
  mutate(number_of_variants_per_gene = n()) %>% 
  distinct(SYMBOL, .keep_all = TRUE) %>% 
  select(-IDs) %>% 
  
  group_by(known_CH_gene) %>% 
  mutate(number_of_gene = n()) %>%
  ungroup() %>% 
  mutate(average_of_variants_per_gene = number_of_variants_per_gene / number_of_gene) %>%
  distinct(known_CH_gene, .keep_all = TRUE) %>% 
  arrange(desc(average_of_variants_per_gene))
```

## Age distribution of known CH variant (can do the same for all potential variants)
```{r}
known_CH_variants <- CH_variants %>% 
  filter(str_detect(SYMBOL, known_CH_gene_list)) %>% 
  distinct(IDs, .keep_all = TRUE) %>% 

  select(-sample_frequency, -sample_count, -nbr_individuals) %>% 

  mutate(age_hist_het_bin_freq = str_match(INFO, "age_hist_het_bin_freq=(.*?);")[,2]) %>%
  # "Count of age values falling below lowest histogram bin edge for heterozygous individuals">
  mutate("<30" = as.numeric(str_match(INFO, "age_hist_het_n_smaller=(.*?);")[,2])) %>%
  # "Count of age values falling above highest histogram bin edge for heterozygous individuals">
  mutate(">80" = as.numeric(str_match(INFO, "age_hist_het_n_larger=(.*?);")[,2])) %>% 
  separate(col = age_hist_het_bin_freq,
           into = c("30-35","35-40","40-45","45-50",
                    "50-55","55-60","60-65","65-70",
                    "70-75","75-80"), 
           sep = "\\|", remove = TRUE, extra = "warn", fill = "right") %>% 
  mutate(across(grep("^[[:digit:]]", colnames(.)), ~ as.numeric(.))) %>% 

  add_row("IDs"="All individuals", "ID"="All individuals of any genotype bin",
          "<30" = 2547, "30-35" = 3423,
          "35-40" = 4546,"40-45" = 8487,
          "45-50" = 10355,"50-55" = 12693,
          "55-60" = 11933,"60-65" = 10534,
          "65-70" = 8882,"70-75" = 5991,
          "75-80" = 4136,
          ">80" = 1935) %>%
  mutate(nbr_individuals = rowSums(select(.,`<30`:`>80`), na.rm = TRUE))

known_CH_variants <- known_CH_variants %>% 
  pivot_longer(cols = c(grep("^([[:digit:]]|<|>)", colnames(.))), 
               names_to = "age_bin", values_to = "sample_count") %>% 
  mutate(age_bin = factor(age_bin, 
                          levels = c("<30", "30-35","35-40","40-45","45-50","50-55",
                                     "55-60","60-65","65-70","70-75","75-80", ">80"
                                     ))) %>% 
  mutate(bin_age = case_when(
    # The average age at diagnosis is 8 overall (ages 0 to 19), 
    # 5 years old for children (aged 0 to 14), and 17 years old 
    # for adolescents (aged 15 to 19), while adults' average 
    # age for cancer diagnosis is 65
    str_detect(age_bin, "<30")              ~ 15, 
    str_detect(age_bin, "30-35")            ~ 32.5,
    str_detect(age_bin, "35-40")            ~ 37.5,
    str_detect(age_bin, "40-45")            ~ 42.5,
    str_detect(age_bin, "45-50")            ~ 47.5,
    str_detect(age_bin, "50-55")            ~ 52.5,
    str_detect(age_bin, "55-60")            ~ 57.5,
    str_detect(age_bin, "60-65")            ~ 62.5,
    str_detect(age_bin, "65-70")            ~ 67.5,
    str_detect(age_bin, "70-75")            ~ 72.5,
    str_detect(age_bin, "75-80")            ~ 77.5,
    TRUE ~ 90
  )) %>% 
  mutate(sample_frequency = sample_count / nbr_individuals)

n <- 0

set.seed(1234)

for(i in unique(known_CH_variants$IDs)) {
  
  All <- known_CH_variants %>% 
    filter(IDs == "All individuals"
  ) %>% 
    mutate(IDs = "Reference")
  data_plot <- known_CH_variants %>% 
    filter(IDs == i
    ) %>% 
    bind_rows(., All) %>%
    distinct(IDs, bin_age, .keep_all = TRUE) # %>% 
    # select(IDs, age_bin, sample_frequency,
    #        bin_age, #distribution_mean,
    #        nbr_individuals, mann_pval, adjusted_pval,
    #        
    #        HGVSP, occurrence_in_cosmic, 
    #        Mutation.somatic.status,
    #        variant_in_cosmic, #prot_change, 
    #        adjusted_pval)
  n <- n + 1
  p <- data_plot %>%
    ggplot(aes(x = bin_age, y= sample_frequency, color= IDs))+
    # ggplot(aes(x = age_bin, y= sample_frequency, color= IDs))+
    # geom_bar(stat = "identity", aes(alpha= IDs))
    geom_smooth(alpha= 0.5, se = FALSE, method = "loess",
                # method.args=list(family=quasibinomial)#,
                span = 0.6
    )+
    labs(title = n)+
    geom_col(data = data_plot %>%
               filter(IDs == i) ,
             aes(x= bin_age, y= sample_frequency),
             position = "identity",
             fill = "red",
             alpha = 0.1)+
    ylim(0, max(data_plot$sample_frequency))
  
  p <- p+
    # geom_vline(xintercept = data_plot$distribution_mean, color= "red")+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("p-MannW= ", round(mann_pval, 3))),
      colour = "black",
      hjust   = 0,
      vjust   = -9
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("BH= ", round(adjusted_pval, 3))),
      colour = "black",
      hjust   = 0,
      vjust   = -11
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("Effect size Hedges g= ", round(effect_size, 3))),
      hjust   = 0,
      vjust   = -17
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("Hedges cat= ", effect_size_cat)),
      hjust   = 0,
      vjust   = -15
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("nbr_individuals= ", nbr_individuals)),
      colour = "darkblue",
      hjust   = 0,
      vjust   = -min(data_plot$sample_frequency)
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste("Is in cosmic?", variant_in_cosmic,",",Mutation.somatic.status)),
      colour = "purple",
      hjust   = 0,
      vjust   = -35
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("HGVSP= ", HGVSP)),
      colour = "purple",
      hjust   = 0,
      vjust   = -33
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("occurrence= ", occurrence_in_cosmic)),
      colour = "purple",
      hjust   = 0,
      vjust   = -31
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("SYMBOL= ", SYMBOL)),
      colour = "blue",
      hjust   = -2,
      vjust   = -29
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene_name_cosmic= ", gene_name_cosmic)),
      colour = "blue",
      hjust   = -1.29,
      vjust   = -25
    )
  
  print(p)

}
```

## VEP : Higher rank only - known CH variant (can do the same for all potential variants)
```{r}
variants_in_gnomAD <- CH_variants %>% 
  distinct(IDs, .keep_all = TRUE) 
variants_in_gnomAD %>% 
  select(Consequence, consequence_rank, known_CH_genes) %>% 
  tbl_summary(by = known_CH_genes) %>% 
  bold_labels() %>% add_overall() %>% 
  add_p() %>% bold_p(t=.05)
```

## VEP : Extract Consequence
I extracted the VEP annotations data from `INFO`.
VEP contains these information :
`"Allele", "Consequence", "IMPACT", "SYMBOL", "Gene", "Feature_type", "Feature", "BIOTYPE", "EXON", "INTRON", "HGVSc", "HGVSp", "`
  `cDNA_position", "CDS_position", "Protein_position", "Amino_acids", "Codons", "Existing_variation", "ALLELE_NUM", "DISTANCE", "`
  `STRAND", "FLAGS", "VARIANT_CLASS", "MINIMISED", "SYMBOL_SOURCE", "HGNC_ID", "CANONICAL", "TSL", "APPRIS", "CCDS", "ENSP", "SWISSPROT", "`
  `TREMBL", "UNIPARC", "GENE_PHENO", "SIFT", "PolyPhen", "DOMAINS", "HGVS_OFFSET", "GMAF", "AFR_MAF", "AMR_MAF", "EAS_MAF", "EUR_MAF", "`
  `SAS_MAF", "AA_MAF", "EA_MAF", "ExAC_MAF", "ExAC_Adj_MAF", "ExAC_AFR_MAF", "ExAC_AMR_MAF", "ExAC_EAS_MAF", "ExAC_FIN_MAF", "`
  `ExAC_NFE_MAF", "ExAC_OTH_MAF", "ExAC_SAS_MAF", "CLIN_SIG", "SOMATIC", "PHENO", "PUBMED", "MOTIF_NAME", "MOTIF_POS", "`
  `HIGH_INF_POS", "MOTIF_SCORE_CHANGE", "LoF", "LoF_filter", "LoF_flags", "LoF_info"`

```{r}
vep_in_gnomAD <- variants_in_gnomAD %>%
  mutate(ens_vep = str_match(INFO, ";vep=(.*?)$")[,2])
vep_in_gnomAD <- vep_in_gnomAD %>%
  separate(col = ens_vep, paste("ens_vep", 1:25, sep=""),
           sep = ",", remove = T, extra = "warn", fill = "right") %>%
  keep(~!all(is.na(.))) %>%
  pivot_longer(cols = starts_with("ens_vep"), names_to = NULL, values_to = "ens_vep") %>%
  drop_na(ens_vep)

ens_vep_var_names <- c("Allele", "Consequence", "IMPACT", "SYMBOL", "Gene", "Feature_type",
                       "Feature", "BIOTYPE", "EXON", "INTRON", "HGVSc", "HGVSp",
                       "cDNA_position", "CDS_position", "Protein_position", "Amino_acids",
                       "Codons", "Existing_variation", "ALLELE_NUM", "DISTANCE", "STRAND",
                       "FLAGS", "VARIANT_CLASS", "MINIMISED", "SYMBOL_SOURCE", "HGNC_ID",
                       "CANONICAL", "TSL", "APPRIS", "CCDS", "ENSP", "SWISSPROT", "TREMBL",
                       "UNIPARC", "GENE_PHENO", "SIFT", "PolyPhen", "DOMAINS", "HGVS_OFFSET",
                       "GMAF", "AFR_MAF", "AMR_MAF", "EAS_MAF", "EUR_MAF", "SAS_MAF", "AA_MAF",
                       "EA_MAF", "ExAC_MAF", "ExAC_Adj_MAF", "ExAC_AFR_MAF", "ExAC_AMR_MAF",
                       "ExAC_EAS_MAF", "ExAC_FIN_MAF", "ExAC_NFE_MAF", "ExAC_OTH_MAF",
                       "ExAC_SAS_MAF", "CLIN_SIG", "SOMATIC", "PHENO", "PUBMED", "MOTIF_NAME",
                       "MOTIF_POS", "HIGH_INF_POS", "MOTIF_SCORE_CHANGE", "LoF", "LoF_filter",
                       "LoF_flags", "LoF_info")

vep_in_gnomAD <- vep_in_gnomAD %>%
  separate(col = ens_vep, into = ens_vep_var_names,
           sep = "\\|", remove = F, extra = "warn", fill = "right")
```

multiple VEP per variants
```{r}
vep_in_gnomAD %>%
  select(IDs, SYMBOL, Consequence, HGVSp,
         "Feature_type", "BIOTYPE", "HGVSc", "HGVSp", "Protein_position", "Amino_acids",
         "VARIANT_CLASS", "SIFT", "PolyPhen",
         "GMAF",
         "AFR_MAF", "AMR_MAF", "EAS_MAF", "EUR_MAF", "SAS_MAF", "AA_MAF", "EA_MAF", "ExAC_MAF",
         "ExAC_Adj_MAF", "ExAC_AFR_MAF", "ExAC_AMR_MAF", "ExAC_EAS_MAF", "ExAC_FIN_MAF",
         "ExAC_NFE_MAF", "ExAC_OTH_MAF", "ExAC_SAS_MAF", "CLIN_SIG", "SOMATIC", "PHENO",
         "MOTIF_NAME", "MOTIF_POS", "HIGH_INF_POS", "MOTIF_SCORE_CHANGE", "LoF",
         "LoF_filter", "LoF_flags", "LoF_info")
```


 - known CH gene
```{r}
vep_in_gnomAD %>%
  filter(str_detect(SYMBOL, known_CH_gene_list)) %>% 
  distinct(IDs, .keep_all = TRUE) %>%
  unite(IDs, c(IDs, SYMBOL)) %>% 
  select(Consequence,
         BIOTYPE,
         HGVSc,
         VARIANT_CLASS,
         GENE_PHENO,
         SIFT,
         PolyPhen,
         CLIN_SIG,
         SOMATIC,
         PHENO) %>%
  tbl_summary() %>%
  bold_labels() %>% 
  modify_header(
    label = "**VEP in gmomAD Characteristic**"
  ) %>%
  modify_spanning_header(all_stat_cols() ~ "**Variants**")
```

## VAF - known CH gene
```{r, fig.width=12, fig.height=25}
variants_in_gnomAD %>% 
  filter(str_detect(SYMBOL, known_CH_gene_list)) %>% 
  select(IDs,
         nc_freq_allele_count, nc_freq_allele_count_female, nc_freq_allele_count_male,
         nc_freq_allele_count_afr, nc_freq_allele_count_afr_female, nc_freq_allele_count_afr_male,
         nc_freq_allele_count_amr, nc_freq_allele_count_amr_female, nc_freq_allele_count_amr_male) %>%
  pivot_longer(cols = -IDs) %>%
  mutate(value = as.numeric(value)) %>%
  mutate(Sex = str_extract(name, "female|male"), Sex = case_when(
    is.na(Sex)             ~ "overall",
    TRUE                   ~ Sex
  )) %>%
  mutate(Ethnicity = str_extract(name, "afr|amr"), Ethnicity = case_when(
    is.na(Ethnicity)       ~ "overall",
    TRUE                   ~ Ethnicity
  )) %>%
  ggplot(aes(x=name, y= value, shape= Sex, color= Ethnicity))+
  geom_point(size= 3)+
  labs(title = "Frequency of the allele in Non-Cancer population")+
  theme_classic(base_size = 16)+
  theme(legend.position = "none")+
  facet_wrap(. ~ IDs, scales = "free", ncol = 2)+
  coord_flip()
```


# Cosmic Analysis for each variants
here 1 time / variant counted
```{r plot variant in cosmic}
CH_variants %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
  ggplot(aes(x=variant_in_cosmic))+
  geom_bar()+
  ggtitle("Number of unique variants in Cosmic")
```

```{r}
variants_in_cosmic <- CH_variants %>% 
  filter(variant_in_cosmic == "Yes")
```
In chromosome `r as.character(variants_in_cosmic %>% distinct(X.CHROM))`, the number of variants present in Cosmic is `r nrow(variants_in_cosmic %>% distinct(IDs))`.

<!-- ### Restricted analysis to few variants -->
<!-- 2-25234373-C-T is the CH known variant in DNMT3A gene + 2-25234374-G-A    -->
<!-- 2-206304912-G-A is in ZDBF2 gene -->
<!-- ```{r variants_in_cosmic} -->
<!-- variants_in_cosmic <- variants_in_cosmic %>%  -->
<!--   filter(str_detect(IDs, "2-25234373-C-T|2-25234374-G-A|2-206304912-G-A")) -->
<!-- rm(mannw_variants) -->
<!-- ``` -->

The variable available in Cosmic are:
`"gene_name_cosmic", "Accession.Number", "Gene.CDS.length", "HGNC.ID", "Sample.name", `
                         `  "ID_sample", "ID_tumour", "Primary.site", "Site.subtype.1", "Site.subtype.2", `
                         `  "Site.subtype.3", "Primary.histology", "Histology.subtype.1", "Histology.subtype.2", `
                        `   "Histology.subtype.3", "Genome.wide.screen", "GENOMIC_MUTATION_ID", `
                        `   "LEGACY_MUTATION_ID", "MUTATION_ID", "Mutation.CDS", "Mutation.AA", `
                         `  "Mutation.Description", "Mutation.zygosity", "LOH", "GRCh", `
                        `   "Mutation.genome.position", "Mutation.strand", "Resistance.Mutation", `
                        `   "FATHMM.prediction", "FATHMM.score", "Mutation.somatic.status", "Pubmed_PMID", `
                        `   "ID_STUDY", "Sample.Type", "Tumour.origin", "Age", "HGVSP", "HGVSC", "HGVSG"`
                        
`r emo::ji("warning")` 1 variant can have multiple samples and primary site, etc
<div class = "row">
<div class = "col-md-6">
```{r, fig.width=12, fig.height=15}
variants_in_cosmic %>% 
  count(SYMBOL) %>% 
  ggplot(aes(x=fct_reorder(SYMBOL, n), y=n, fill= str_detect(SYMBOL, known_CH_gene_list)))+
  geom_bar(stat = "identity")+
  ggtitle("Number of overall samples per gene")+
  scale_fill_discrete(name = NULL, labels = c("Not CH genes", "CH genes"))+
  labs(x= NULL, y="Number of samples")+
  coord_flip()
```
</div>

<div class = "col-md-6">
```{r, fig.width=12, fig.height=15}
variants_in_cosmic %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
  count(SYMBOL) %>% 
  ggplot(aes(x=fct_reorder(SYMBOL, n), y=n, fill= str_detect(SYMBOL, known_CH_gene_list)))+
  geom_bar(stat = "identity")+
  ggtitle("Number of variants per gene")+
  scale_fill_discrete(name = NULL, labels = c("Not CH genes", "CH genes"))+
  coord_flip()
```
</div>
</div>




## Cosmic Patients/Cell Line Characteristics

```{r Cosmic Patients/Cell Line Characteristics}
# show_header_names(tbl)

variants_in_cosmic %>% 
  unite(IDs, c(IDs, SYMBOL)) %>% 
  select(Age,
         Primary.site, Primary.histology, 
         Mutation.Description,
         FATHMM.prediction, Mutation.somatic.status) %>% 
  mutate(Primary.site = factor(Primary.site, ordered = FALSE),
         Primary.site = relevel(Primary.site, ref = "haematopoietic_and_lymphoid_tissue")
           ) %>%
  mutate(Primary.histology = factor(Primary.histology, ordered = FALSE),
         Primary.histology = relevel(Primary.histology, ref = "haematopoietic_neoplasm")
           ) %>%
  tbl_summary() %>% 
  bold_labels() %>% 
  modify_header(
    label = "**Cosmic Patients/Cell Line Characteristics**"
  ) %>%
  modify_spanning_header(all_stat_cols() ~ "**Variants**")
```

<!-- ### each sample of a variant is counted. -->
<!-- ```{r sample plot, fig.height=20} -->
<!-- cosmic_var1 <- c("Age", -->
<!--          "Primary.site", "Primary.histology", "Tumour.origin", -->
<!--          "Sample.Type", -->
<!--          "Mutation.Description", -->
<!--          "FATHMM.prediction", "Mutation.somatic.status")  -->

<!-- for (i in cosmic_var1){ -->
<!--   print( -->
<!--     variants_in_cosmic %>% -->
<!--       filter(str_detect(SYMBOL, known_CH_gene_list)) %>%  -->
<!--       ggplot(aes(x= .data[[i]] ))+ -->
<!--       geom_bar()+ -->
<!--       facet_wrap(.~ IDs, ncol = 3) -->
<!--   ) -->
<!-- } -->
<!-- ``` -->

<!-- ### 1 variant is counted once -->
<!-- ```{r variant plot} -->
<!-- for (i in cosmic_var1){ -->
<!--   print( -->
<!--     variants_in_cosmic %>% -->
<!--       filter(str_detect(SYMBOL, known_CH_gene_list)) %>%  -->
<!--       distinct(IDs, .data[[i]]) %>%  -->
<!--       ggplot(aes(x= .data[[i]] ))+ -->
<!--       geom_bar() -->
<!--   ) -->
<!-- } -->
<!-- # # bar_plot <- function(data){ -->
<!-- #   for (i in cosmic_var1){ -->
<!-- #     # if(class(markers_ROI[[i]]) == "factor" | class(markers_ROI[[i]]) == "character") { -->
<!-- #       variants_in_cosmic %>% -->
<!-- #         ggplot()+ -->
<!-- #         geom_bar(aes(x=.data[[i]])) -->
<!-- #       # p2 <- markers_ROI %>% -->
<!-- #       #   ggplot()+ -->
<!-- #       #   geom_violin(aes(x=.data[[i]], y= .data[[paste0(type, ".p")]]), scale = "count")+ -->
<!-- #       #   geom_boxplot(aes(x=.data[[i]], y= .data[[paste0(type, ".p")]]), width=.1) + -->
<!-- #       #   theme_classic2()+ -->
<!-- #       #   labs(x=NULL, y=NULL, title="Peripheral \nROIs") -->
<!-- #       # main <- str_match(type, "percent_(.*)_total")[,2] -->
<!-- #       # gridExtra::grid.arrange(p1, p2, ncol = 2, -->
<!-- #       #                         top = "% of Cells Present in", -->
<!-- #       #                         left =  paste("Overall", main, "immune cells (%) per patients"), -->
<!-- #       #                         bottom = "Data on Race-Matched Patients") -->
<!-- #     # } -->
<!-- #   } -->
<!-- # # } -->
<!-- # bar_plot(data = variants_in_cosmic) -->
<!-- ``` -->




<!-- ## Cosmic Gene Characteristics -->
```{r Cosmic Gene Characteristics}
# variants_in_cosmic %>% 
#   unite(IDs, c(IDs, SYMBOL, gene_name_cosmic, Mutation.AA)) %>% 
#   select(#gene_name_cosmic,
#          Mutation.Description, Mutation.zygosity,
#          FATHMM.prediction, FATHMM.score, Mutation.somatic.status,
#          IDs) %>% 
#   tbl_summary(by= IDs) %>% 
#   bold_labels() %>% add_overall() %>% 
#   modify_header(
#     label = '**Cosmic Gene Characteristics**'
#   ) %>%
#   modify_spanning_header(all_stat_cols() ~ "**Variants**")
```

<!-- ### each sample of a variant is counted. -->
<!-- ```{r cosmic plot, fig.height=20} -->
<!-- cosmic_var2 <- c("gene_name_cosmic", -->
<!--          "Mutation.AA",  -->
<!--          "Mutation.Description", -->
<!--          "FATHMM.prediction", "FATHMM.score", "Mutation.somatic.status")  -->

<!-- for (i in cosmic_var2){ -->
<!--   print( -->
<!--     variants_in_cosmic %>% -->
<!--       filter(str_detect(SYMBOL, known_CH_gene_list)) %>%  -->
<!--       ggplot(aes(x= .data[[i]] ))+ -->
<!--       geom_bar()+ -->
<!--       facet_wrap(.~ IDs, ncol = 2) -->
<!--   ) -->
<!-- } -->
<!-- ``` -->

<!-- ### 1 variant is counted once -->
<!-- ```{r} -->
<!-- for (i in cosmic_var2){ -->
<!--   print( -->
<!--     variants_in_cosmic %>% -->
<!--       filter(str_detect(SYMBOL, known_CH_gene_list)) %>%  -->
<!--       distinct(IDs, .data[[i]]) %>%  -->
<!--       ggplot(aes(x= .data[[i]] ))+ -->
<!--       geom_bar() -->
<!--   ) -->
<!-- } -->

<!-- ``` -->


















# Bick Analysis for each variants
<!-- **CH R882H mutation is depicted in red**   -->
<!-- bh adjusted_selected in yellow    -->
<!-- effect size_selected in orange -->
```{r}
R882H <- "2-25234373-C-T"
bhadjusted_selected <- paste0(c("2-25244379-G-C",
                                "2-25244494-C-T", "2-25248052-G-A", "2-25274935-C-T",
                                "2-25282744-G-A"),
                              collapse = "|")
effsize_selected <- paste0(c("2-25234321-G-A", "2-25234374-G-A",
                                 "2-25328714-C-G", "2-25328736-G-A"),
                              collapse = "|")
```

# Definition

Mutation somatic status :  
Information on whether the sample was reported to be Confirmed Somatic, Previously Reported or Variant of unknown origin  
- Previously observed = when the mutation has been reported as somatic previously but not in current paper.  
- Confirmed Somatic = if the mutation has been confimed to be somatic in the experiment by sequencing both the tumour and a matched normal from the same patient.  
- Variant of unknown origin = when the mutation is known to be somatic but the tumour was sequenced without a matched normal.


INFO=<ID=AC,Number=A,Type=Integer,Description="Alternate allele count for samples  
afr, African-American/African ancestry
amr, Latino ancestry
asj, Ashkenazi Jewish ancestry
eas, East Asian ancestry  
fin, Finnish ancestry  
nfe, Non-Finnish European ancestry  
nfe_bgr, Bulgarian (Eastern European) ancestry  
nfe_est, Estonian ancestry  
nfe_nwe, North-Western European ancestry  
nfe_onf, Other Non-Finnish European ancestry  
nfe_seu, Southern European ancestry  
nfe_swe, Swedish ancestry  
oth, Other ancestry  

chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://docs.gdc.cancer.gov/Data/PDF/Data_UG.pdf

GMAF Non-reference allele and frequency of existing variant in 1000 Genomes   
ExAC_MAF Frequency of existing variant in ExAC combined population  
ExAC_NFE_MAF Frequency of existing variant in ExAC Non-Finnish European population   
ExAC_OTH_MAF Frequency of existing variant in ExAC combined other combined populations   
ExAC_SAS_MAF Frequency of existing variant in ExAC South Asian population   
CLIN_SIG Clinical significance of variant from dbSNP Somatic status of existing variant(s)  
SOMATIC 
PHENO Indicates if existing variant is associated with a phenotype, disease or trait  





