---
title: "Inverstigate chr2 pval"
author: "Christelle Colin-Leitzinger"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_collapsed: no
    theme: cerulean
    highlight: pygments
    df_print: paged
editor_options:
  chunk_output_type: console
---

<style type="text/css">
h1.title {
  font-size: 25px;
}
.figure {
   margin-top: 25px;
   margin-bottom: 25px;
}

table {
    margin-top: 25px;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      fig.align='center'
                      )
options(gtsummary.print_engine = "gt")
options(gtsummary.as_gt.addl_cmds = "gt::tab_options(table.font.size = 12, data_row.padding = gt::px(1))")
```


<style>
div.blue { background-color:#0099CC; border-radius: 5px; padding: 20px; font-size: 38px}
</style>
<div class = "blue">

<span style="color: white;">GnomAD data</span>

</div>
<br>

```{r library, include=FALSE}
library(tidyverse)
library(effsize)

theme_set(theme_classic())
```

```{r load}
path <- fs::path("", "Volumes", "Lab_Gillis", "Christelle")
# gnomad_pval <-
#   read.delim(paste0(path, "/gnomAD_raw_data/chr2_test/padjust_age_selected_variants/chr2_padjusted.vcf.gz"),
#               sep = " ")
# gnomad_pval <-
#   read.delim(paste0(path, "/gnomAD_raw_data/splitted_data/age_comparison_pval/padjust_age_selected_variants/chr2_padjusted.vcf.gz"),
#               sep = " ")
gnomad <-
  read.delim(paste0(path, "/gnomAD_raw_data/splitted_data/age_comparison_pval/padjust_age_selected_variants/chr2_padjusted.vcf.gz"),
              sep = " ")

```

```{r}
# gnomad <- gnomad %>% 
#   left_join(., gnomad_pval %>% 
#               select(IDs, adjusted_pval),
#             by = "IDs")

# rm(gnomad_pval)
```

<br>

***

<br>

## `r emo::ji("spiral_notepad")` Definition
Mutation somatic status :  
Information on whether the sample was reported to be Confirmed Somatic, Previously Reported or Variant of unknown origin  
- Previously observed = when the mutation has been reported as somatic previously but not in current paper.  
- Confirmed Somatic = if the mutation has been confimed to be somatic in the experiment by sequencing both the tumour and a matched normal from the same patient.  
- Variant of unknown origin = when the mutation is known to be somatic but the tumour was sequenced without a matched normal.


-- Found in INFO variable:  

- Race/ethnicity (AC and AF available for each)  
AC,Description="Alternate allele count for samples"  
afr, African-American/African ancestry  
amr, Latino ancestry  
asj, Ashkenazi Jewish ancestry  
eas, East Asian ancestry  
sas, South Asian ancestry  
fin, Finnish ancestry  
nfe, Non-Finnish European ancestry  
nfe_bgr, Bulgarian (Eastern European) ancestry  
nfe_est, Estonian ancestry  
nfe_nwe, North-Western European ancestry  
nfe_onf, Other Non-Finnish European ancestry  
nfe_seu, Southern European ancestry  
nfe_swe, Swedish ancestry  
oth, Other ancestry  

AC_popmax, Description="Allele count in the population with the maximum AF"

- Sex (AC and AF available for each)  
AC_female, Description="Alternate allele count for female samples"  
AC_male, Description="Alternate allele count for male samples"  

- Age
age_hist_het_bin_freq, Description="Histogram of ages of heterozygous individuals; bin edges are: 30.0|35.0|40.0|45.0|50.0|55.0|60.0|65.0|70.0|75.0|80.0; total number of individuals of any genotype bin: 2547|3423|4546|8487|10355|12693|11933|10534|8882|5991|4136|1935"  
age_hist_het_n_larger, Description="Count of age values falling above highest histogram bin edge for heterozygous individuals"  
age_hist_het_n_smaller, Description="Count of age values falling below lowest histogram bin edge for heterozygous individuals"  
age_hist_hom_bin_freq, Description="Histogram of ages of homozygous alternate individuals; bin edges are: 30.0|35.0|40.0|45.0|50.0|55.0|60.0|65.0|70.0|75.0|80.0; total number of individuals of any genotype bin: 2547|3423|4546|8487|10355|12693|11933|10534|8882|5991|4136|1935"  
age_hist_hom_n_larger, Description="Count of age values falling above highest histogram bin edge for homozygous alternate individuals"  
age_hist_hom_n_smaller, Description="Count of age values falling below lowest histogram bin edge for homozygous alternate individuals"  

- Allele/Variant info
allele_type, Description="Allele type (snv, ins, del, or mixed)"  
variant_type, Description="Variant type (snv, indel, multi-snv, multi-indel, or mixed)"  
vep, Description="Consequence annotations from Ensembl VEP.  
Format: Allele|Consequence|IMPACT|SYMBOL|Gene|Feature_type|Feature|BIOTYPE|EXON|INTRON|HGVSc|HGVSp|cDNA_position|CDS_position|Protein_position|Amino_acids|Codons|Existing_variation|ALLELE_NUM|DISTANCE|STRAND|FLAGS|VARIANT_CLASS|MINIMISED|SYMBOL_SOURCE|HGNC_ID|CANONICAL|TSL|APPRIS|CCDS|ENSP|SWISSPROT|TREMBL|UNIPARC|GENE_PHENO|SIFT|PolyPhen|DOMAINS|HGVS_OFFSET|GMAF|AFR_MAF|AMR_MAF|EAS_MAF|EUR_MAF|SAS_MAF|AA_MAF|EA_MAF|ExAC_MAF|ExAC_Adj_MAF|ExAC_AFR_MAF|ExAC_AMR_MAF|ExAC_EAS_MAF|ExAC_FIN_MAF|ExAC_NFE_MAF|ExAC_OTH_MAF|ExAC_SAS_MAF|CLIN_SIG|SOMATIC|PHENO|PUBMED|MOTIF_NAME|MOTIF_POS|HIGH_INF_POS|MOTIF_SCORE_CHANGE|LoF|LoF_filter|LoF_flags|LoF_info"  
was_mixed, Description="Variant type was mixed">

- Population (AC and AF available for each)  
gnomAD v2.1.1 
gnomAD v2.1.1 (non-TOPMed): Only samples that are not present in the Trans-Omics for Precision Medicine (TOPMed)/BRAVO release. The allele counts in this subset can thus be added to those of BRAVO to federate both datasets.  
gnomAD v2.1.1 (non-cancer): Only samples from individuals who were not ascertained for having cancer in a cancer study.  
gnomAD v2.1.1 (non-neuro): Only samples from individuals who were not ascertained for having a neurological condition in a neurological case/control study.  
gnomAD v2.1.1 (controls): Only samples from individuals who were not selected as a case in a case/control study of common disease.  

<span style="color:red">**So we don't have general/precise number/info on race/eth, sex, pop for each variant**</span>

<br>

***

<br>

# I. Mean and effect size
**CH R882H mutation is depicted in red**  
bh adjusted_selected in yellow   
effect size_selected in orange
```{r plot}
bhadjusted_selected <- paste0(c("2-25244379-G-C", 
                                "2-25244494-C-T", "2-25248052-G-A", "2-25274935-C-T", 
                                "2-25282744-G-A"),
                              collapse = "|")
effsize_selected <- paste0(c("2-25234321-G-A", "2-25234374-G-A",
                                 "2-25328714-C-G", "2-25328736-G-A"),
                              collapse = "|")

gnomad %>% 
  select(IDs, nbr_individuals, mann_w, adjusted_pval) %>% 
  ggplot(aes(x= adjusted_pval, y= mann_w))+
  geom_point(color= "grey")+
  geom_point(data= gnomad %>% filter(IDs=="2-25234373-C-T"),
             aes(x= adjusted_pval, y=mann_w), color= "red")+
  geom_point(data= gnomad %>% 
               filter(str_detect(IDs, bhadjusted_selected)),
             aes(x= adjusted_pval, y=mann_w), color= "yellow")+
  geom_point(data= gnomad %>% 
               filter(str_detect(IDs, effsize_selected)),
             aes(x= adjusted_pval, y=mann_w), color= "orange")+
  theme(legend.position = "none")

gnomad %>% 
  select(IDs, nbr_individuals, mann_w, adjusted_pval) %>% 
  ggplot(aes(x= mann_w, y= nbr_individuals))+
  geom_point(color= "grey")+
  geom_point(data= gnomad %>% filter(IDs=="2-25234373-C-T"),
             aes(x= mann_w, y=adjusted_pval), color= "red")+
  geom_point(data= gnomad %>% 
               filter(str_detect(IDs, bhadjusted_selected)),
             aes(x= mann_w, y=adjusted_pval), color= "yellow")+
  geom_point(data= gnomad %>% 
               filter(str_detect(IDs, effsize_selected)),
             aes(x= mann_w, y=adjusted_pval), color= "orange")+
  theme(legend.position = "none")

gnomad %>% 
  select(IDs, nbr_individuals, mann_w, adjusted_pval) %>% 
  ggplot(aes(x= adjusted_pval, y= nbr_individuals))+
  geom_point(color= "grey")+
  geom_point(data= gnomad %>% filter(IDs=="2-25234373-C-T"),
             aes(x= adjusted_pval, y=nbr_individuals), color= "red")+
  geom_point(data= gnomad %>% 
               filter(str_detect(IDs, bhadjusted_selected)),
             aes(x= adjusted_pval, y=nbr_individuals), color= "yellow")+
  geom_point(data= gnomad %>% 
               filter(str_detect(IDs, effsize_selected)),
             aes(x= adjusted_pval, y=nbr_individuals), color= "orange")+
  theme(legend.position = "none")
```

```{r plot1}
# All <- gnomad %>% 
#   filter(IDs == "All individuals"
#   ) %>% 
#   # mutate(IDs = "Reference") %>% 
#   select(distribution_mean_ref = distribution_mean)
# 
# data_plot <- gnomad %>%
#   # filter(IDs == i
#   # ) %>% 
#   select("IDs", distribution_mean, adjusted_pval) %>% 
#   bind_cols(., All) %>% 
#   mutate(effect = distribution_mean - distribution_mean_ref) # need to keep sup values

# data_plot %>% 
#   slice(1:20) %>% 
#   ggplot(aes(x = IDs, y= effect, color= effect > 0
#              #, color= IDs == "All individuals"
#   ))+
#   geom_point()+
#   geom_hline(yintercept = 0)+
#   coord_flip()
```

<!-- On all the variants -->
```{r plot3}
library(ggforce)
# data_plot %>% 
#   ggplot(aes(x= effect, y= adjusted_pval))+
#   geom_point(color= "grey")+
#   geom_point(data= data_plot %>% filter(IDs=="2-25234373-C-T"),
#              aes(x= effect, y=adjusted_pval), color= "red")+
#   geom_point(data= data_plot %>% 
#                filter(str_detect(IDs, bhadjusted_selected)),
#              aes(x= effect, y=adjusted_pval), color= "yellow")+
#   geom_point(data= data_plot %>% 
#                filter(str_detect(IDs, effsize_selected)),
#              aes(x= effect, y=adjusted_pval), color= "orange")+
#   scale_y_log10()+
#   labs(y= "BH pval (log10)")+
#   theme(legend.position = "none")
# 
# data_plot %>% 
#   ggplot(aes(x= effect, y= adjusted_pval))+
#   geom_point(color= "grey")+
#   geom_point(data= data_plot %>% filter(IDs=="2-25234373-C-T"),
#              aes(x= effect, y=adjusted_pval), color= "red")+
#   geom_point(data= data_plot %>% 
#                filter(str_detect(IDs, bhadjusted_selected)),
#              aes(x= effect, y=adjusted_pval), color= "yellow")+
#   geom_point(data= data_plot %>% 
#                filter(str_detect(IDs, effsize_selected)),
#              aes(x= effect, y=adjusted_pval), color= "orange")+
#   scale_y_continuous(trans = trans_reverser('log10'))+
#   # scale_y_continuous(trans = "log2")+
#   labs(y= "BH pval (log10)")+
#   theme(legend.position = "none")
```

<!-- On BH < 0.05 -->
```{r plot4}
library(ggforce)
# data_plot %>% 
#   filter(adjusted_pval < 0.05) %>% 
#   ggplot(aes(x= effect, y= adjusted_pval))+
#   geom_point(color= "grey")+
#   geom_point(data= data_plot %>% filter(IDs=="2-25234373-C-T"),
#              aes(x= effect, y=adjusted_pval), color= "red")+
#   geom_point(data= data_plot %>% 
#                filter(str_detect(IDs, bhadjusted_selected)),
#              aes(x= effect, y=adjusted_pval), color= "yellow")+
#   geom_point(data= data_plot %>% 
#                filter(str_detect(IDs, effsize_selected)),
#              aes(x= effect, y=adjusted_pval), color= "orange")+
#   scale_y_log10()+
#   labs(y= "BH pval (log10)")+
#   theme(legend.position = "none")
# 
# data_plot %>% 
#   filter(adjusted_pval < 0.05) %>% 
#   ggplot(aes(x= effect, y= adjusted_pval))+
#   geom_point(color= "grey")+
#   geom_point(data= data_plot %>% filter(IDs=="2-25234373-C-T"),
#              aes(x= effect, y=adjusted_pval), color= "red")+
#   geom_point(data= data_plot %>% 
#                filter(str_detect(IDs, bhadjusted_selected)),
#              aes(x= effect, y=adjusted_pval), color= "yellow")+
#   geom_point(data= data_plot %>% 
#                filter(str_detect(IDs, effsize_selected)),
#              aes(x= effect, y=adjusted_pval), color= "orange")+
#   scale_y_continuous(trans = 'log2')+
#   # scale_y_continuous(trans = "log2")+
#   labs(y= "BH pval (log2)")+
#   theme(legend.position = "none")
```

|**Mutation Consequence in gnomAD**                                                                                           | **N = 12,003,450** |
|:----------------------------------------------------------------------------------------------------------------------------|:------------------:|
|__Consequence__                                                                                                              |                    |
|3_prime_UTR_variant                                                                                                          |   159,202 (1.3%)   |
|3_prime_UTR_variant&NMD_transcript_variant                                                                                   |   187,791 (1.6%)   |
|5_prime_UTR_variant                                                                                                          |   134,118 (1.1%)   |
|5_prime_UTR_variant&NMD_transcript_variant                                                                                   |   13,401 (0.1%)    |
|coding_sequence_variant                                                                                                      |    214 (<0.1%)     |
|coding_sequence_variant&3_prime_UTR_variant                                                                                  |    176 (<0.1%)     |
|coding_sequence_variant&3_prime_UTR_variant&NMD_transcript_variant                                                           |     17 (<0.1%)     |
|coding_sequence_variant&5_prime_UTR_variant                                                                                  |    102 (<0.1%)     |
|coding_sequence_variant&5_prime_UTR_variant&NMD_transcript_variant                                                           |     28 (<0.1%)     |
|coding_sequence_variant&NMD_transcript_variant                                                                               |     87 (<0.1%)     |
|downstream_gene_variant                                                                                                      |  2,293,818 (19%)   |
|frameshift_variant                                                                                                           |   28,582 (0.2%)    |
|frameshift_variant&NMD_transcript_variant                                                                                    |   3,203 (<0.1%)    |
|frameshift_variant&splice_region_variant                                                                                     |   1,479 (<0.1%)    |
|frameshift_variant&splice_region_variant&intron_variant                                                                      |     26 (<0.1%)     |
|frameshift_variant&splice_region_variant&NMD_transcript_variant                                                              |     96 (<0.1%)     |
|frameshift_variant&splice_region_variant&stop_retained_variant                                                               |     1 (<0.1%)      |
|frameshift_variant&splice_region_variant&stop_retained_variant&NMD_transcript_variant                                        |     1 (<0.1%)      |
|frameshift_variant&start_lost                                                                                                |     83 (<0.1%)     |
|frameshift_variant&start_lost&NMD_transcript_variant                                                                         |     20 (<0.1%)     |
|frameshift_variant&stop_lost                                                                                                 |    196 (<0.1%)     |
|frameshift_variant&stop_lost&NMD_transcript_variant                                                                          |     17 (<0.1%)     |
|frameshift_variant&stop_lost&splice_region_variant                                                                           |     5 (<0.1%)      |
|frameshift_variant&stop_lost&splice_region_variant&NMD_transcript_variant                                                    |     1 (<0.1%)      |
|frameshift_variant&stop_retained_variant                                                                                     |     70 (<0.1%)     |
|frameshift_variant&stop_retained_variant&NMD_transcript_variant                                                              |     4 (<0.1%)      |
|incomplete_terminal_codon_variant                                                                                            |     1 (<0.1%)      |
|incomplete_terminal_codon_variant&coding_sequence_variant                                                                    |    372 (<0.1%)     |
|incomplete_terminal_codon_variant&coding_sequence_variant&3_prime_UTR_variant                                                |     13 (<0.1%)     |
|inframe_deletion                                                                                                             |   24,388 (0.2%)    |
|inframe_deletion&NMD_transcript_variant                                                                                      |   2,606 (<0.1%)    |
|inframe_deletion&splice_region_variant                                                                                       |    340 (<0.1%)     |
|inframe_deletion&splice_region_variant&NMD_transcript_variant                                                                |     20 (<0.1%)     |
|inframe_insertion                                                                                                            |   8,525 (<0.1%)    |
|inframe_insertion&coding_sequence_variant                                                                                    |     1 (<0.1%)      |
|inframe_insertion&NMD_transcript_variant                                                                                     |    712 (<0.1%)     |
|inframe_insertion&splice_region_variant                                                                                      |    226 (<0.1%)     |
|inframe_insertion&splice_region_variant&NMD_transcript_variant                                                               |     16 (<0.1%)     |
|inframe_insertion&stop_retained_variant                                                                                      |     6 (<0.1%)      |
|intergenic_variant                                                                                                           |   1,235 (<0.1%)    |
|intron_variant                                                                                                               |  2,411,266 (20%)   |
|intron_variant&NMD_transcript_variant                                                                                        |   377,703 (3.1%)   |
|intron_variant&non_coding_transcript_variant                                                                                 |   812,342 (6.8%)   |
|mature_miRNA_variant                                                                                                         |    503 (<0.1%)     |
|missense_variant                                                                                                             |  1,298,410 (11%)   |
|missense_variant&NMD_transcript_variant                                                                                      |   83,818 (0.7%)    |
|missense_variant&splice_region_variant                                                                                       |   35,430 (0.3%)    |
|missense_variant&splice_region_variant&NMD_transcript_variant                                                                |   2,612 (<0.1%)    |
|non_coding_transcript_exon_variant&non_coding_transcript_variant                                                             |   789,417 (6.6%)   |
|non_coding_transcript_variant                                                                                                |     19 (<0.1%)     |
|protein_altering_variant                                                                                                     |    218 (<0.1%)     |
|protein_altering_variant&incomplete_terminal_codon_variant                                                                   |     2 (<0.1%)      |
|protein_altering_variant&NMD_transcript_variant                                                                              |     69 (<0.1%)     |
|protein_altering_variant&splice_region_variant                                                                               |     30 (<0.1%)     |
|protein_altering_variant&splice_region_variant&NMD_transcript_variant                                                        |     3 (<0.1%)      |
|regulatory_region_variant                                                                                                    |   379,867 (3.2%)   |
|splice_acceptor_variant                                                                                                      |   5,933 (<0.1%)    |
|splice_acceptor_variant&3_prime_UTR_variant                                                                                  |     2 (<0.1%)      |
|splice_acceptor_variant&3_prime_UTR_variant&intron_variant                                                                   |     5 (<0.1%)      |
|splice_acceptor_variant&3_prime_UTR_variant&intron_variant&NMD_transcript_variant                                            |     46 (<0.1%)     |
|splice_acceptor_variant&3_prime_UTR_variant&NMD_transcript_variant                                                           |     39 (<0.1%)     |
|splice_acceptor_variant&5_prime_UTR_variant                                                                                  |     35 (<0.1%)     |
|splice_acceptor_variant&5_prime_UTR_variant&intron_variant                                                                   |     28 (<0.1%)     |
|splice_acceptor_variant&5_prime_UTR_variant&intron_variant&NMD_transcript_variant                                            |     1 (<0.1%)      |
|splice_acceptor_variant&5_prime_UTR_variant&NMD_transcript_variant                                                           |     2 (<0.1%)      |
|splice_acceptor_variant&coding_sequence_variant                                                                              |    274 (<0.1%)     |
|splice_acceptor_variant&coding_sequence_variant&3_prime_UTR_variant                                                          |     2 (<0.1%)      |
|splice_acceptor_variant&coding_sequence_variant&3_prime_UTR_variant&intron_variant                                           |     15 (<0.1%)     |
|splice_acceptor_variant&coding_sequence_variant&3_prime_UTR_variant&intron_variant&NMD_transcript_variant                    |     1 (<0.1%)      |
|splice_acceptor_variant&coding_sequence_variant&5_prime_UTR_variant&intron_variant                                           |     12 (<0.1%)     |
|splice_acceptor_variant&coding_sequence_variant&5_prime_UTR_variant&intron_variant&NMD_transcript_variant                    |     4 (<0.1%)      |
|splice_acceptor_variant&coding_sequence_variant&intron_variant                                                               |    539 (<0.1%)     |
|splice_acceptor_variant&coding_sequence_variant&intron_variant&NMD_transcript_variant                                        |     42 (<0.1%)     |
|splice_acceptor_variant&coding_sequence_variant&NMD_transcript_variant                                                       |     31 (<0.1%)     |
|splice_acceptor_variant&frameshift_variant                                                                                   |     4 (<0.1%)      |
|splice_acceptor_variant&intron_variant                                                                                       |    367 (<0.1%)     |
|splice_acceptor_variant&intron_variant&NMD_transcript_variant                                                                |     55 (<0.1%)     |
|splice_acceptor_variant&intron_variant&non_coding_transcript_variant                                                         |     86 (<0.1%)     |
|splice_acceptor_variant&NMD_transcript_variant                                                                               |   1,110 (<0.1%)    |
|splice_acceptor_variant&non_coding_transcript_exon_variant&intron_variant&non_coding_transcript_variant                      |    131 (<0.1%)     |
|splice_acceptor_variant&non_coding_transcript_exon_variant&non_coding_transcript_variant                                     |     93 (<0.1%)     |
|splice_acceptor_variant&non_coding_transcript_variant                                                                        |   1,752 (<0.1%)    |
|splice_acceptor_variant&splice_donor_variant&coding_sequence_variant&intron_variant                                          |     10 (<0.1%)     |
|splice_acceptor_variant&splice_donor_variant&inframe_deletion&intron_variant                                                 |     1 (<0.1%)      |
|splice_acceptor_variant&splice_donor_variant&intron_variant                                                                  |     6 (<0.1%)      |
|splice_acceptor_variant&splice_donor_variant&intron_variant&non_coding_transcript_variant                                    |     2 (<0.1%)      |
|splice_acceptor_variant&splice_donor_variant&non_coding_transcript_exon_variant&intron_variant&non_coding_transcript_variant |     5 (<0.1%)      |
|splice_acceptor_variant&synonymous_variant                                                                                   |     1 (<0.1%)      |
|splice_donor_variant                                                                                                         |   8,050 (<0.1%)    |
|splice_donor_variant&3_prime_UTR_variant                                                                                     |     2 (<0.1%)      |
|splice_donor_variant&3_prime_UTR_variant&intron_variant                                                                      |     15 (<0.1%)     |
|splice_donor_variant&3_prime_UTR_variant&intron_variant&NMD_transcript_variant                                               |     31 (<0.1%)     |
|splice_donor_variant&3_prime_UTR_variant&NMD_transcript_variant                                                              |     16 (<0.1%)     |
|splice_donor_variant&5_prime_UTR_variant                                                                                     |     6 (<0.1%)      |
|splice_donor_variant&5_prime_UTR_variant&intron_variant                                                                      |     9 (<0.1%)      |
|splice_donor_variant&5_prime_UTR_variant&intron_variant&NMD_transcript_variant                                               |     2 (<0.1%)      |
|splice_donor_variant&coding_sequence_variant                                                                                 |    212 (<0.1%)     |
|splice_donor_variant&coding_sequence_variant&3_prime_UTR_variant                                                             |     2 (<0.1%)      |
|splice_donor_variant&coding_sequence_variant&3_prime_UTR_variant&intron_variant                                              |     2 (<0.1%)      |
|splice_donor_variant&coding_sequence_variant&3_prime_UTR_variant&intron_variant&NMD_transcript_variant                       |     2 (<0.1%)      |
|splice_donor_variant&coding_sequence_variant&5_prime_UTR_variant                                                             |     3 (<0.1%)      |
|splice_donor_variant&coding_sequence_variant&5_prime_UTR_variant&intron_variant                                              |     3 (<0.1%)      |
|splice_donor_variant&coding_sequence_variant&intron_variant                                                                  |    451 (<0.1%)     |
|splice_donor_variant&coding_sequence_variant&intron_variant&NMD_transcript_variant                                           |     31 (<0.1%)     |
|splice_donor_variant&coding_sequence_variant&NMD_transcript_variant                                                          |     20 (<0.1%)     |
|splice_donor_variant&frameshift_variant                                                                                      |     3 (<0.1%)      |
|splice_donor_variant&intron_variant                                                                                          |    400 (<0.1%)     |
|splice_donor_variant&intron_variant&NMD_transcript_variant                                                                   |     68 (<0.1%)     |
|splice_donor_variant&intron_variant&non_coding_transcript_variant                                                            |    109 (<0.1%)     |
|splice_donor_variant&NMD_transcript_variant                                                                                  |   1,421 (<0.1%)    |
|splice_donor_variant&non_coding_transcript_exon_variant&intron_variant&non_coding_transcript_variant                         |    154 (<0.1%)     |
|splice_donor_variant&non_coding_transcript_exon_variant&non_coding_transcript_variant                                        |     57 (<0.1%)     |
|splice_donor_variant&non_coding_transcript_variant                                                                           |   2,434 (<0.1%)    |
|splice_donor_variant&splice_acceptor_variant&3_prime_UTR_variant&intron_variant&NMD_transcript_variant                       |     1 (<0.1%)      |
|splice_donor_variant&splice_acceptor_variant&coding_sequence_variant&intron_variant                                          |     3 (<0.1%)      |
|splice_donor_variant&splice_acceptor_variant&frameshift_variant&intron_variant                                               |     2 (<0.1%)      |
|splice_donor_variant&splice_acceptor_variant&inframe_deletion&intron_variant                                                 |     3 (<0.1%)      |
|splice_donor_variant&splice_acceptor_variant&intron_variant                                                                  |     3 (<0.1%)      |
|splice_donor_variant&splice_acceptor_variant&non_coding_transcript_exon_variant&intron_variant&non_coding_transcript_variant |     1 (<0.1%)      |
|splice_region_variant&3_prime_UTR_variant                                                                                    |    634 (<0.1%)     |
|splice_region_variant&3_prime_UTR_variant&NMD_transcript_variant                                                             |   5,430 (<0.1%)    |
|splice_region_variant&5_prime_UTR_variant                                                                                    |   3,395 (<0.1%)    |
|splice_region_variant&5_prime_UTR_variant&NMD_transcript_variant                                                             |    177 (<0.1%)     |
|splice_region_variant&coding_sequence_variant&5_prime_UTR_variant                                                            |     2 (<0.1%)      |
|splice_region_variant&incomplete_terminal_codon_variant&coding_sequence_variant                                              |     3 (<0.1%)      |
|splice_region_variant&intron_variant                                                                                         |   188,765 (1.6%)   |
|splice_region_variant&intron_variant&NMD_transcript_variant                                                                  |   28,145 (0.2%)    |
|splice_region_variant&intron_variant&non_coding_transcript_variant                                                           |   49,130 (0.4%)    |
|splice_region_variant&non_coding_transcript_exon_variant&non_coding_transcript_variant                                       |   18,152 (0.2%)    |
|splice_region_variant&non_coding_transcript_variant                                                                          |    164 (<0.1%)     |
|splice_region_variant&stop_retained_variant                                                                                  |     4 (<0.1%)      |
|splice_region_variant&stop_retained_variant&NMD_transcript_variant                                                           |     4 (<0.1%)      |
|splice_region_variant&synonymous_variant                                                                                     |   25,010 (0.2%)    |
|splice_region_variant&synonymous_variant&NMD_transcript_variant                                                              |   1,551 (<0.1%)    |
|start_lost                                                                                                                   |   2,527 (<0.1%)    |
|start_lost&inframe_deletion                                                                                                  |     15 (<0.1%)     |
|start_lost&inframe_deletion&NMD_transcript_variant                                                                           |     11 (<0.1%)     |
|start_lost&NMD_transcript_variant                                                                                            |    326 (<0.1%)     |
|start_lost&splice_region_variant                                                                                             |     69 (<0.1%)     |
|start_lost&splice_region_variant&NMD_transcript_variant                                                                      |     7 (<0.1%)      |
|stop_gained                                                                                                                  |   19,602 (0.2%)    |
|stop_gained&frameshift_variant                                                                                               |    429 (<0.1%)     |
|stop_gained&frameshift_variant&NMD_transcript_variant                                                                        |     35 (<0.1%)     |
|stop_gained&frameshift_variant&splice_region_variant                                                                         |     60 (<0.1%)     |
|stop_gained&frameshift_variant&splice_region_variant&NMD_transcript_variant                                                  |     3 (<0.1%)      |
|stop_gained&inframe_deletion                                                                                                 |     8 (<0.1%)      |
|stop_gained&inframe_deletion&splice_region_variant                                                                           |     1 (<0.1%)      |
|stop_gained&inframe_insertion                                                                                                |     69 (<0.1%)     |
|stop_gained&inframe_insertion&NMD_transcript_variant                                                                         |     2 (<0.1%)      |
|stop_gained&inframe_insertion&splice_region_variant                                                                          |     8 (<0.1%)      |
|stop_gained&inframe_insertion&splice_region_variant&NMD_transcript_variant                                                   |     1 (<0.1%)      |
|stop_gained&NMD_transcript_variant                                                                                           |   1,863 (<0.1%)    |
|stop_gained&protein_altering_variant                                                                                         |     19 (<0.1%)     |
|stop_gained&protein_altering_variant&splice_region_variant                                                                   |     5 (<0.1%)      |
|stop_gained&splice_region_variant                                                                                            |    545 (<0.1%)     |
|stop_gained&splice_region_variant&NMD_transcript_variant                                                                     |     63 (<0.1%)     |
|stop_lost                                                                                                                    |   1,076 (<0.1%)    |
|stop_lost&inframe_deletion                                                                                                   |     5 (<0.1%)      |
|stop_lost&inframe_deletion&NMD_transcript_variant                                                                            |     2 (<0.1%)      |
|stop_lost&inframe_deletion&splice_region_variant                                                                             |     7 (<0.1%)      |
|stop_lost&NMD_transcript_variant                                                                                             |    260 (<0.1%)     |
|stop_lost&splice_region_variant                                                                                              |     8 (<0.1%)      |
|stop_lost&splice_region_variant&NMD_transcript_variant                                                                       |     18 (<0.1%)     |
|stop_retained_variant                                                                                                        |    709 (<0.1%)     |
|stop_retained_variant&NMD_transcript_variant                                                                                 |    120 (<0.1%)     |
|synonymous_variant                                                                                                           |   933,604 (7.8%)   |
|synonymous_variant&NMD_transcript_variant                                                                                    |   51,541 (0.4%)    |
|TF_binding_site_variant                                                                                                      |   15,611 (0.1%)    |
|TF_binding_site_variant&TFBS_ablation                                                                                        |    112 (<0.1%)     |
|TFBS_ablation&TF_binding_site_variant                                                                                        |    115 (<0.1%)     |
|upstream_gene_variant                                                                                                        |  1,576,970 (13%)   |


|**Mutation Consequence for Cosmic Patients/Cell Line**                                 | **N = 988,048** |
|:--------------------------------------------------------------------------------------|:---------------:|
|__Consequence__                                                                        |                 |
|3_prime_UTR_variant                                                                    |  10,918 (1.1%)  |
|3_prime_UTR_variant&NMD_transcript_variant                                             |  22,747 (2.3%)  |
|5_prime_UTR_variant                                                                    |  10,115 (1.0%)  |
|5_prime_UTR_variant&NMD_transcript_variant                                             |   716 (<0.1%)   |
|coding_sequence_variant                                                                |   16 (<0.1%)    |
|coding_sequence_variant&NMD_transcript_variant                                         |    8 (<0.1%)    |
|downstream_gene_variant                                                                |  175,178 (18%)  |
|incomplete_terminal_codon_variant&coding_sequence_variant                              |   42 (<0.1%)    |
|intergenic_variant                                                                     |    8 (<0.1%)    |
|intron_variant                                                                         |  115,260 (12%)  |
|intron_variant&NMD_transcript_variant                                                  |  17,909 (1.8%)  |
|intron_variant&non_coding_transcript_variant                                           |  48,186 (4.9%)  |
|mature_miRNA_variant                                                                   |   27 (<0.1%)    |
|missense_variant                                                                       |  190,694 (19%)  |
|missense_variant&NMD_transcript_variant                                                |  14,115 (1.4%)  |
|missense_variant&splice_region_variant                                                 |  4,286 (0.4%)   |
|missense_variant&splice_region_variant&NMD_transcript_variant                          |   286 (<0.1%)   |
|non_coding_transcript_exon_variant&non_coding_transcript_variant                       |  80,736 (8.2%)  |
|regulatory_region_variant                                                              |  29,654 (3.0%)  |
|splice_acceptor_variant                                                                |   327 (<0.1%)   |
|splice_acceptor_variant&NMD_transcript_variant                                         |   50 (<0.1%)    |
|splice_acceptor_variant&non_coding_transcript_variant                                  |   69 (<0.1%)    |
|splice_donor_variant                                                                   |   729 (<0.1%)   |
|splice_donor_variant&NMD_transcript_variant                                            |   138 (<0.1%)   |
|splice_donor_variant&non_coding_transcript_variant                                     |   167 (<0.1%)   |
|splice_region_variant&3_prime_UTR_variant                                              |   51 (<0.1%)    |
|splice_region_variant&3_prime_UTR_variant&NMD_transcript_variant                       |   566 (<0.1%)   |
|splice_region_variant&5_prime_UTR_variant                                              |   294 (<0.1%)   |
|splice_region_variant&5_prime_UTR_variant&NMD_transcript_variant                       |   12 (<0.1%)    |
|splice_region_variant&intron_variant                                                   |  7,469 (0.8%)   |
|splice_region_variant&intron_variant&NMD_transcript_variant                            |  1,019 (0.1%)   |
|splice_region_variant&intron_variant&non_coding_transcript_variant                     |  1,770 (0.2%)   |
|splice_region_variant&non_coding_transcript_exon_variant&non_coding_transcript_variant |  1,806 (0.2%)   |
|splice_region_variant&synonymous_variant                                               |  2,943 (0.3%)   |
|splice_region_variant&synonymous_variant&NMD_transcript_variant                        |   179 (<0.1%)   |
|start_lost                                                                             |   122 (<0.1%)   |
|start_lost&NMD_transcript_variant                                                      |   19 (<0.1%)    |
|start_lost&splice_region_variant                                                       |    6 (<0.1%)    |
|start_lost&splice_region_variant&NMD_transcript_variant                                |    1 (<0.1%)    |
|stop_gained                                                                            |  2,951 (0.3%)   |
|stop_gained&NMD_transcript_variant                                                     |   265 (<0.1%)   |
|stop_gained&splice_region_variant                                                      |   77 (<0.1%)    |
|stop_gained&splice_region_variant&NMD_transcript_variant                               |    7 (<0.1%)    |
|stop_lost                                                                              |   104 (<0.1%)   |
|stop_lost&NMD_transcript_variant                                                       |   13 (<0.1%)    |
|stop_lost&splice_region_variant                                                        |    1 (<0.1%)    |
|stop_lost&splice_region_variant&NMD_transcript_variant                                 |    3 (<0.1%)    |
|stop_retained_variant                                                                  |   51 (<0.1%)    |
|stop_retained_variant&NMD_transcript_variant                                           |   12 (<0.1%)    |
|synonymous_variant                                                                     |  121,768 (12%)  |
|synonymous_variant&NMD_transcript_variant                                              |  6,928 (0.7%)   |
|TF_binding_site_variant                                                                |   892 (<0.1%)   |
|upstream_gene_variant                                                                  |  116,338 (12%)  |



```{r recreate age bin}
gnomad_decoded <- gnomad %>%
  select(-age_bin, -sample_frequency, -sample_count, -nbr_individuals) %>% 
  mutate(gene_in_gnomad = str_match(INFO, "(DNMT3A)(.*?)$")[,2]) %>% 
  filter(IDs == "2-25457242-C-T" | IDs == "2-25234373-C-T" | gene_in_gnomad == "DNMT3A") %>%
  # filter(INFO == "Arg882His" | gene_in_gnomad == "DNMT3A") %>% 
  # mutate(prot_change = str_match(INFO, "(|R)(.*?)|$")[,2]) %>% 
  # mutate(prot_change1 = str_match(INFO, "(|R)(.*?)|$")[,3]) %>% 
  # slice(1:50) %>% 
  # mutate(IDs = ifelse(IDs == "2-25234373-C-T", "2-25234373-C-T_R882H", IDs)) %>% 

  mutate(age_hist_het_bin_freq = str_match(INFO, "age_hist_het_bin_freq=(.*?);")[,2]) %>%
  # "Count of age values falling below lowest histogram bin edge for heterozygous individuals">
  mutate("<30_het" = as.numeric(str_match(INFO, "age_hist_het_n_smaller=(.*?);")[,2])) %>%
  # "Count of age values falling above highest histogram bin edge for heterozygous individuals">
  mutate(">80_het" = as.numeric(str_match(INFO, "age_hist_het_n_larger=(.*?);")[,2])) %>% 
  separate(col = age_hist_het_bin_freq,
           into = c("30-35_het","35-40_het","40-45_het","45-50_het",
                    "50-55_het","55-60_het","60-65_het","65-70_het",
                    "70-75_het","75-80_het"), 
           sep = "\\|", remove = TRUE, extra = "warn", fill = "right") %>% 
  mutate(across(grep("^[[:digit:]]", colnames(.)), ~ as.numeric(.))) %>% 

  # 2. Extract count per age bin for heterozygous ind--
  mutate(age_hist_hom_bin_freq = str_match(INFO, "age_hist_hom_bin_freq=(.*?);")[,2]) %>%
  # "Count of age values falling below lowest histogram bin edge for homozygous individuals">
  mutate("<30_hom" = as.numeric(str_match(INFO, "age_hist_hom_n_smaller=(.*?);")[,2])) %>%
  # "Count of age values falling above highest histogram bin edge for homozygous individuals">
  mutate(">80_hom" = as.numeric(str_match(INFO, "age_hist_hom_n_larger=(.*?);")[,2])) %>% 
  separate(col = age_hist_hom_bin_freq,
           into = c("30-35_hom","35-40_hom","40-45_hom","45-50_hom",
                    "50-55_hom","55-60_hom","60-65_hom","65-70_hom",
                    "70-75_hom","75-80_hom"), 
           sep = "\\|", remove = TRUE, extra = "warn", fill = "right") %>% 
  mutate(across(grep("^[[:digit:]]", colnames(.)), ~ as.numeric(.))) %>% 
  
  # 3. Combine count for Het and hom individuals
  mutate("<30" = rowSums(select(.,`30-35_het`,`30-35_hom`), na.rm = TRUE),
         "30-35" = rowSums(select(.,`30-35_het`,`30-35_hom`), na.rm = TRUE),
         "35-40" = rowSums(select(.,`35-40_het`,`35-40_hom`), na.rm = TRUE),
         "40-45" = rowSums(select(.,`40-45_het`,`40-45_hom`), na.rm = TRUE),
         "45-50" = rowSums(select(.,`45-50_het`,`45-50_hom`), na.rm = TRUE),
         "50-55" = rowSums(select(.,`50-55_het`,`50-55_hom`), na.rm = TRUE),
         "55-60" = rowSums(select(.,`55-60_het`,`55-60_hom`), na.rm = TRUE),
         "60-65" = rowSums(select(.,`60-65_het`,`60-65_hom`), na.rm = TRUE),
         "65-70" = rowSums(select(.,`65-70_het`,`65-70_hom`), na.rm = TRUE),
         "70-75" = rowSums(select(.,`70-75_het`,`70-75_hom`), na.rm = TRUE),
         "75-80" = rowSums(select(.,`75-80_het`,`75-80_hom`), na.rm = TRUE),
         ">80" = rowSums(select(.,`>80_het`,`>80_hom`), na.rm = TRUE)
         ) %>%
  add_row("IDs"="All individuals", "ID"="All individuals of any genotype bin",
          "<30" = 2547, "30-35" = 3423,
          "35-40" = 4546,"40-45" = 8487,
          "45-50" = 10355,"50-55" = 12693,
          "55-60" = 11933,"60-65" = 10534,
          "65-70" = 8882,"70-75" = 5991,
          "75-80" = 4136,
          ">80" = 1935) %>%
  mutate(nbr_individuals = rowSums(select(.,`<30`:`>80`), na.rm = TRUE)) %>%
  
  select(-c(ends_with("_het"), ends_with("_hom"), bin_age)) %>% 
  
  select(IDs, everything())
```

```{r recreate age bin2}
gnomad_decoded <- gnomad_decoded %>% 
  pivot_longer(cols = c(grep("[[:digit:]]", colnames(.))), 
               names_to = "age_bin", values_to = "sample_count") %>% 
  mutate(age_bin = factor(age_bin, 
                          levels = c("<30", "30-35","35-40","40-45","45-50","50-55",
                                     "55-60","60-65","65-70","70-75","75-80", ">80"
                                     ))) %>% 
  mutate(bin_age = case_when(
    # The average age at diagnosis is 8 overall (ages 0 to 19), 
    # 5 years old for children (aged 0 to 14), and 17 years old 
    # for adolescents (aged 15 to 19), while adults' average 
    # age for cancer diagnosis is 65
    str_detect(age_bin, "<30")              ~ 15, 
    str_detect(age_bin, "30-35")            ~ 32.5,
    str_detect(age_bin, "35-40")            ~ 37.5,
    str_detect(age_bin, "40-45")            ~ 42.5,
    str_detect(age_bin, "45-50")            ~ 47.5,
    str_detect(age_bin, "50-55")            ~ 52.5,
    str_detect(age_bin, "55-60")            ~ 57.5,
    str_detect(age_bin, "60-65")            ~ 62.5,
    str_detect(age_bin, "65-70")            ~ 67.5,
    str_detect(age_bin, "70-75")            ~ 72.5,
    str_detect(age_bin, "75-80")            ~ 77.5,
    TRUE ~ 90
  )) %>% 
  mutate(sample_frequency = sample_count / nbr_individuals)
```

<!-- <!-- Found in Cosmic data --> -->

```{r load cosmic}
# cosmic <-
#   read.delim(paste0(path, "/cosmic_raw_data/cosmic_DNMT3_data.vcf.gz"),
#              header = FALSE,
#              col.names = c("Gene.name", "Accession.Number", "Gene.CDS.length", "HGNC.ID", "Sample.name",
#                            "ID_sample", "ID_tumour", "Primary.site", "Site.subtype.1", "Site.subtype.2",
#                            "Site.subtype.3", "Primary.histology", "Histology.subtype.1", "Histology.subtype.2",
#                            "Histology.subtype.3", "Genome.wide.screen", "GENOMIC_MUTATION_ID",
#                            "LEGACY_MUTATION_ID", "MUTATION_ID", "Mutation.CDS", "Mutation.AA",
#                            "Mutation.Description", "Mutation.zygosity", "LOH", "GRCh",
#                            "Mutation.genome.position", "Mutation.strand", "Resistance.Mutation",
#                            "FATHMM.prediction", "FATHMM.score", "Mutation.somatic.status", "Pubmed_PMID",
#                            "ID_STUDY", "Sample.Type", "Tumour.origin", "Age", "HGVSP", "HGVSC", "HGVSG"))

cosmic <-
  read.delim(paste0(path, "/cosmic_raw_data/splitted_data/filter_cosmic_withIDs/chr2_cosmic.vcf.gz"),
              sep = " ")
```

```{r}
# cosmic <-
#   cosmic %>%
#   mutate(modification = str_match(HGVSG, "(delins|ins|del|dup)(.*?)$")[,2],
#          aa_inserted = case_when(
#            modification %in% c("ins", "delins")   ~ str_match(HGVSG, "(delins|ins|del|dup)(.*?)$")[,3],
#            TRUE                                   ~ NA_character_
#          )) %>%
# 
#   separate(HGVSG, into = c("chr", "chr_arm", "POS", "ALT"), sep = ":|\\.|>|delins|dup|del", remove = FALSE) %>%
#   mutate(REF = str_match(POS, "([0-9]*)([A-Z]*)")[,3]) %>%
#   mutate(POS = str_match(POS, "([0-9]*)([A-Z]*)")[,2],
#          POS = as.numeric(POS)) %>%
#   # separate(Mutation.genome.position, into = c("chr", "start", "end"), sep = ":|-", remove = FALSE) %>%
#   unite(IDs, c(chr, POS, REF, ALT), sep = "-", remove = FALSE) %>% #select(IDs, everything())
#   # separate(HGVSG, into = c("ENSP", "p_prot_modif"), sep = ":", remove = FALSE) %>%
#   mutate(IDs = case_when(
#     IDs == "-NA-NA-NA"                 ~ NA_character_,
#     TRUE                               ~ IDs
#   )) %>%
#   separate(`Gene.name`, into = c("gene_name_cosmic"), sep = "_", remove = FALSE) %>%
#   # 1 tumor can have multiple sample
#   distinct(IDs, ID_tumour, .keep_all = TRUE) %>%
#   arrange(IDs) %>%
#   select(IDs, "Gene.name", gene_name_cosmic, Gene.name, "Accession.Number", "chr", "POS", REF, "ALT", HGVSG, everything())
```

```{r}
# cosmic <- cosmic %>%
#   group_by(IDs, gene_name_cosmic, chr, POS, REF, ALT) %>%
#   mutate(occurrence_in_cosmic = n())

# max_occurence = max(cosmic$occurrence_in_cosmic)
# cosmic1 <- cosmic %>% ###################################### Clean up here--------------------------------
# 
#   # distinct()
#   group_by(IDs, gene_name_cosmic, chr, POS, REF, ALT,
#            HGNC.ID,
#            HGVSP, ###################################### Clean up here--------------------------------
#            FATHMM.prediction, FATHMM.score,
# 
#            occurrence_in_cosmic) %>%
#   summarise_at(vars(Accession.Number, HGVSG,
#                     Gene.CDS.length,
#     Mutation.CDS,
#           Mutation.AA,
#           Primary.site,
#           Mutation.Description,
#     Mutation.genome.position,
#           Primary.histology,
#           Histology.subtype.1, Histology.subtype.2, Histology.subtype.3,
#           HGVSC,
#           Mutation.somatic.status, Pubmed_PMID,
#           Sample.Type, Tumour.origin, Age), str_c, collapse = "; ") %>%
#   ungroup() %>%
#   separate(Age, paste("Age", 1:max(max_occurence), sep = "_"),
#            sep = "; ", remove = TRUE,
#            extra = "warn", fill = "right") %>%
#   purrr::keep(~!all(is.na(.))) %>%
#   mutate(across(starts_with("Age_"), ~ as.numeric(.))) %>%
# 
#   mutate(average_age = rowMeans(select(., starts_with("Age")), na.rm = TRUE)) %>%
#   mutate(variant_in_cosmic = "Yes")
```

```{r bind cosmic Ids}
gnomad_decoded <- left_join(gnomad_decoded %>%
                              mutate(X.CHROM = as.character(X.CHROM)),
                            cosmic %>%
                              mutate(chr = as.character(chr)),
                            by = c("IDs", "X.CHROM" = "chr", "POS", "REF", "ALT"))
```

# II. Plots all variants
**R882H is curve #6 aka 2-25234373-C-T.**
```{r Plots all variants}
# mann_w <-  data.frame(matrix(nrow=1, ncol=0)) 
effect_size <-  data.frame(matrix(nrow=1, ncol=0)) 
effect_size_cat <-  data.frame(matrix(nrow=1, ncol=0)) 

n <- 0

set.seed(1234)

for(i in unique(gnomad_decoded$IDs)) {
  
  All <- gnomad_decoded %>% 
    filter(IDs == "All individuals"
  ) %>% 
    mutate(IDs = "Reference")
  data_plot <- gnomad_decoded %>% 
    filter(IDs == i
    ) %>% 
    bind_rows(., All) %>%
    distinct(IDs, bin_age, .keep_all = TRUE) # %>% 
    # select(IDs, age_bin, sample_frequency,
    #        bin_age, #distribution_mean,
    #        nbr_individuals, mann_pval, adjusted_pval,
    #        
    #        HGVSP, occurrence_in_cosmic, 
    #        Mutation.somatic.status,
    #        variant_in_cosmic, #prot_change, 
    #        adjusted_pval)
  n <- n + 1
  p <- data_plot %>%
    ggplot(aes(x = bin_age, y= sample_frequency, color= IDs))+
    # ggplot(aes(x = age_bin, y= sample_frequency, color= IDs))+
    # geom_bar(stat = "identity", aes(alpha= IDs))
    geom_smooth(alpha= 0.5, se = FALSE, method = "loess",
                # method.args=list(family=quasibinomial)#,
                span = 0.6
    )+
    labs(title = n)+
    geom_col(data = data_plot %>%
               filter(IDs == i) ,
             aes(x= bin_age, y= sample_frequency),
             position = "identity",
             fill = "red",
             alpha = 0.1)+
    ylim(0, max(data_plot$sample_frequency))
  # 
  # p_ <- layer_data(p, 1) %>% 
  # mutate(colour =  as.factor(colour)) %>% 
  # mutate_if(is.numeric, ~replace(., is.na(.), 0))
  # 
  # variant <- p_ %>% filter(group == 1)
  # variant <- sample(variant$x, 180, prob = variant$y, replace = TRUE)
  # ref <- p_ %>% filter(group == 2)
  # ref <- sample(ref$x, 180, prob = ref$y, replace = TRUE)
  # 
  # # result1 <- wilcox.test(variant, ref, alternative = "greater")$p.value
  # # mann_w <- rbind(mann_w, result1)
  # 
  # result2 <- cohen.d(variant, ref, hedges.correction=TRUE)$estimate
  # effect_size <- rbind(effect_size, result2)
  # result3 <- cohen.d(variant, ref,hedges.correction=TRUE)$magnitude[1] %>% 
  #   as_tibble() %>%
  #   mutate(value = case_when(
  #     value == "negligible"   ~ "negligible",
  #     value == "small"        ~ "small",
  #     value == "medium"       ~ "medium",
  #     value == "large"        ~ "large"
  #   ))
  # effect_size_cat <- rbind(effect_size_cat, result3)
  
  p <- p+
    # geom_vline(xintercept = data_plot$distribution_mean, color= "red")+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("MannW= ", round(mann_pval, 3))),
      colour = "black",
      hjust   = 0,
      vjust   = -9
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("BH= ", round(adjusted_pval, 3))),
      colour = "black",
      hjust   = 0,
      vjust   = -11
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("Effect size Hedges g= ", round(effect_size, 3))),
      hjust   = 0,
      vjust   = -17
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("Hedges cat= ", effect_size_cat)),
      hjust   = 0,
      vjust   = -15
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("nbr_individuals= ", nbr_individuals)),
      colour = "darkblue",
      hjust   = 0,
      vjust   = -min(data_plot$sample_frequency)
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste("Is in cosmic?", variant_in_cosmic,",",Mutation.somatic.status)),
      colour = "purple",
      hjust   = 0,
      vjust   = -35
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("HGVSP= ", HGVSP)),
      colour = "purple",
      hjust   = 0,
      vjust   = -33
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("occurrence= ", occurrence_in_cosmic)),
      colour = "purple",
      hjust   = 0,
      vjust   = -31
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene.name= ", Gene.name)),
      colour = "blue",
      hjust   = -2,
      vjust   = -29
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene_in_gnomad= ", gene_in_gnomad)),
      colour = "blue",
      hjust   = -1.15,
      vjust   = -27
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene_name_cosmic= ", gene_name_cosmic)),
      colour = "blue",
      hjust   = -1.29,
      vjust   = -25
    )
  
  print(p)

}
```


```{r}
# mann_w <- mann_w %>% `colnames<-`(c("mann_w")) %>% 
#   mutate(IDs = c(unique(gnomad_decoded$IDs)))
# effect_size <- effect_size %>% `colnames<-`(c("effect_size"))%>% 
#   mutate(IDs = c(unique(gnomad_decoded$IDs)))
# effect_size_cat <- effect_size_cat %>% `colnames<-`(c("effect_size_cat"))
# 
# stat_measure <- bind_cols(effect_size, effect_size_cat) %>% 
#   left_join(gnomad_decoded, .,
#   by= "IDs")
stat_measure <- gnomad_decoded
# print(head(stat_measure))
```

```{r}
stat_measure %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
  ggplot(aes(x= effect_size, y= adjusted_pval))+
  geom_point(color= "grey")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>% 
               filter(IDs=="2-25234373-C-T"),
             aes(x= effect_size, y=adjusted_pval), color= "red")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>% 
               filter(str_detect(IDs, bhadjusted_selected)),
             aes(x= effect_size, y=adjusted_pval), color= "yellow")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>% 
               filter(str_detect(IDs, effsize_selected)),
             aes(x= effect_size, y=adjusted_pval), color= "orange")

stat_measure %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
  ggplot(aes(x= effect_size, y= adjusted_pval))+
  geom_point(color= "grey")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>% 
               filter(IDs=="2-25234373-C-T"),
             aes(x= effect_size, y=adjusted_pval), color= "red")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>% 
               filter(str_detect(IDs, bhadjusted_selected)),
             aes(x= effect_size, y=adjusted_pval), color= "yellow")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>% 
               filter(str_detect(IDs, effsize_selected)),
             aes(x= effect_size, y=adjusted_pval), color= "orange")+
  scale_y_log10()+
  labs(x= "Hedges g effect size", y= "BH pval (log10)")+
  theme(legend.position = "none")

stat_measure %>% 
  distinct(IDs, .keep_all = TRUE) %>% 
  ggplot(aes(x= effect_size, y= adjusted_pval))+
  geom_point(color= "grey")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>%
               filter(IDs=="2-25234373-C-T"),
             aes(x= effect_size, y=adjusted_pval), color= "red")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>% 
               filter(str_detect(IDs, bhadjusted_selected)),
             aes(x= effect_size, y=adjusted_pval), color= "yellow")+
  geom_point(data= stat_measure %>% 
               distinct(IDs, .keep_all = TRUE) %>% 
               filter(str_detect(IDs, effsize_selected)),
             aes(x= effect_size, y=adjusted_pval), color= "orange")+
  scale_y_continuous(trans = 'log2')+
  # scale_y_continuous(trans = "log2")+
  labs(x= "Hedges g effect size", y= "BH pval (log2)")+
  theme(legend.position = "none")
```

# III. Plots for BH  0.1
**R882H is curve #2 aka 2-25234373-C-T.**
```{r}
selected_variants <- stat_measure %>% filter(adjusted_pval <= 0.1)
n <- 0

set.seed(1234)

for(i in unique(selected_variants$IDs)) {
  
  All <- gnomad_decoded %>% 
    filter(IDs == "All individuals"
  ) %>% 
    mutate(IDs = "Reference")
  data_plot <- selected_variants %>% 
    filter(IDs == i
    ) %>% 
    bind_rows(., All) %>%
    distinct(IDs, bin_age, .keep_all = TRUE) # %>% 
    # select(IDs, age_bin, sample_frequency,
    #        bin_age, #distribution_mean,
    #        nbr_individuals, mann_pval, adjusted_pval,
    #        effect_size, effect_size_cat,
    #        HGVSP, occurrence_in_cosmic, 
    #        Mutation.somatic.status,
    #        variant_in_cosmic, #prot_change, 
    #        adjusted_pval)
  n <- n + 1
  p <- data_plot %>% 
    ggplot(aes(x = bin_age, y= sample_frequency, color= IDs))+
    # ggplot(aes(x = age_bin, y= sample_frequency, color= IDs))+
    # geom_bar(stat = "identity", aes(alpha= IDs))
    geom_smooth(alpha= 0.5, se = FALSE, method = "loess",
                # method.args=list(family=quasibinomial)#,
                span = 0.6
    )+
    labs(title = n)+
    geom_col(data = data_plot %>% 
               filter(IDs == i) ,
             aes(x= bin_age, y= sample_frequency),
             position = "identity",
             fill = "red",
             alpha = 0.1)+
    ylim(0, max(data_plot$sample_frequency))
  
  p <- p+
    # geom_vline(xintercept = data_plot$distribution_mean, color= "red")+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("MannW= ", round(mann_pval, 3))),
      colour = "black",
      hjust   = 0,
      vjust   = -9
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("BH= ", round(adjusted_pval, 3))),
      colour = "black",
      hjust   = 0,
      vjust   = -11
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("Effect size Hedges g= ", round(effect_size, 3))),
      hjust   = 0,
      vjust   = -17
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("Hedges cat= ", effect_size_cat)),
      hjust   = 0,
      vjust   = -15
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste("Is in cosmic?", variant_in_cosmic,",",Mutation.somatic.status)),
      colour = "purple",
      hjust   = 0,
      vjust   = -35
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("HGVSP= ", HGVSP)),
      colour = "purple",
      hjust   = 0,
      vjust   = -33
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("occurrence= ", occurrence_in_cosmic)),
      colour = "purple",
      hjust   = 0,
      vjust   = -31
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene.name= ", Gene.name)),
      colour = "blue",
      hjust   = -2,
      vjust   = -29
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene_in_gnomad= ", gene_in_gnomad)),
      colour = "blue",
      hjust   = -1.15,
      vjust   = -27
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene_name_cosmic= ", gene_name_cosmic)),
      colour = "blue",
      hjust   = -1.29,
      vjust   = -25
    )
  
  print(p)

}
```

# IV. Plots Hedges > small
**R882H is curve #2 aka 2-25234373-C-T.**
```{r}
selected_variants <- selected_variants %>% filter(!str_detect(effect_size_cat, "negligible|small"))
n <- 0

set.seed(1234)

for(i in unique(selected_variants$IDs)) {
  
  All <- gnomad_decoded %>% 
    filter(IDs == "All individuals"
  ) %>% 
    mutate(IDs = "Reference")
  data_plot <- selected_variants %>% 
    filter(IDs == i
    ) %>% 
    bind_rows(., All) %>%
    distinct(IDs, bin_age, .keep_all = TRUE) #%>% 
    # select(IDs, age_bin, sample_frequency,
    #        bin_age, #distribution_mean,
    #        nbr_individuals, mann_pval, adjusted_pval,
    #        effect_size, effect_size_cat,
    #        HGVSP, occurrence_in_cosmic, 
    #        Mutation.somatic.status,
    #        variant_in_cosmic, #prot_change, 
    #        adjusted_pval)
  n <- n + 1
  p <- data_plot %>% 
    ggplot(aes(x = bin_age, y= sample_frequency, color= IDs))+
    # ggplot(aes(x = age_bin, y= sample_frequency, color= IDs))+
    # geom_bar(stat = "identity", aes(alpha= IDs))
    geom_smooth(alpha= 0.5, se = FALSE, method = "loess",
                # method.args=list(family=quasibinomial)#,
                span = 0.6
    )+
    labs(title = n)+
    geom_col(data = data_plot %>% 
               filter(IDs == i) ,
             aes(x= bin_age, y= sample_frequency),
             position = "identity",
             fill = "red",
             alpha = 0.1)+
    ylim(0, max(data_plot$sample_frequency))
  
  p <- p+
    # geom_vline(xintercept = data_plot$distribution_mean, color= "red")+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("MannW= ", round(mann_pval, 3))),
      colour = "black",
      hjust   = 0,
      vjust   = -9
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("BH= ", round(adjusted_pval, 3))),
      colour = "black",
      hjust   = 0,
      vjust   = -11
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("Effect size Hedges g= ", round(effect_size, 3))),
      hjust   = 0,
      vjust   = -17
    )+
    geom_text(
      data    = data_plot %>% 
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("Hedges cat= ", effect_size_cat)),
      hjust   = 0,
      vjust   = -15
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste("Is in cosmic?", variant_in_cosmic,",",Mutation.somatic.status)),
      colour = "purple",
      hjust   = 0,
      vjust   = -35
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("HGVSP= ", HGVSP)),
      colour = "purple",
      hjust   = 0,
      vjust   = -33
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("occurrence= ", occurrence_in_cosmic)),
      colour = "purple",
      hjust   = 0,
      vjust   = -31
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene.name= ", Gene.name)),
      colour = "blue",
      hjust   = -2,
      vjust   = -29
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene_in_gnomad= ", gene_in_gnomad)),
      colour = "blue",
      hjust   = -1.15,
      vjust   = -27
    )+
    geom_text(
      data    = data_plot %>%
               filter(IDs == i),
      mapping = aes(x = -Inf, y = -Inf, label = paste0("gene_name_cosmic= ", gene_name_cosmic)),
      colour = "blue",
      hjust   = -1.29,
      vjust   = -25
    )
  
  print(p)

}
```


# V. Analysis

## 1. Gene
```{r}
stat_measure %>% 
  select(IDs, Sample.name, occurrence_in_cosmic)

stat_measure %>% 
  filter(variant_in_cosmic == "Yes") %>% 
  select(IDs, Sample.name, occurrence_in_cosmic)

stat_measure %>% 
  select(Gene.name, gene_in_gnomad, gene_name_cosmic)

stat_measure %>% 
  select(gene_in_gnomad, INFO) %>% 
  slice(1)
```

## 2. VEP for 1st variant
```{r}
stat_measure1 <- stat_measure %>%
  slice(1)
```

```{r}
vep_in_gnomAD <- stat_measure1 %>%
  mutate(ens_vep = str_match(INFO, ";vep=(.*?)$")[,2])
vep_in_gnomAD <- vep_in_gnomAD %>%
  separate(col = ens_vep, paste("ens_vep", 1:25, sep=""),
           sep = ",", remove = T, extra = "warn", fill = "right") %>%
  keep(~!all(is.na(.))) %>%
  pivot_longer(cols = starts_with("ens_vep"), names_to = NULL, values_to = "ens_vep") %>%
  drop_na(ens_vep)

vep_in_gnomAD %>% 
  select(gene_in_gnomad, ens_vep) 
```


```{r}
ens_vep_var_names <- c("Allele", "Consequence", "IMPACT", "SYMBOL", "Gene", "Feature_type",
                       "Feature", "BIOTYPE", "EXON", "INTRON", "HGVSc", "HGVSp",
                       "cDNA_position", "CDS_position", "Protein_position", "Amino_acids",
                       "Codons", "Existing_variation", "ALLELE_NUM", "DISTANCE", "STRAND",
                       "FLAGS", "VARIANT_CLASS", "MINIMISED", "SYMBOL_SOURCE", "HGNC_ID",
                       "CANONICAL", "TSL", "APPRIS", "CCDS", "ENSP", "SWISSPROT", "TREMBL",
                       "UNIPARC", "GENE_PHENO", "SIFT", "PolyPhen", "DOMAINS", "HGVS_OFFSET",
                       "GMAF", "AFR_MAF", "AMR_MAF", "EAS_MAF", "EUR_MAF", "SAS_MAF", "AA_MAF",
                       "EA_MAF", "ExAC_MAF", "ExAC_Adj_MAF", "ExAC_AFR_MAF", "ExAC_AMR_MAF",
                       "ExAC_EAS_MAF", "ExAC_FIN_MAF", "ExAC_NFE_MAF", "ExAC_OTH_MAF",
                       "ExAC_SAS_MAF", "CLIN_SIG", "SOMATIC", "PHENO", "PUBMED", "MOTIF_NAME",
                       "MOTIF_POS", "HIGH_INF_POS", "MOTIF_SCORE_CHANGE", "LoF", "LoF_filter",
                       "LoF_flags", "LoF_info")

vep_in_gnomAD <- vep_in_gnomAD %>%
  separate(col = ens_vep, into = ens_vep_var_names,
           sep = "\\|", remove = F, extra = "warn", fill = "right")

vep_in_gnomAD %>%
  select(IDs, SYMBOL, Consequence, HGVSp,
         "Feature_type", "BIOTYPE", "HGVSc", "HGVSp", "Protein_position", "Amino_acids",
         "VARIANT_CLASS", "SIFT", "PolyPhen",
         "GMAF",
         "AFR_MAF", "AMR_MAF", "EAS_MAF", "EUR_MAF", "SAS_MAF", "AA_MAF", "EA_MAF", "ExAC_MAF",
         "ExAC_Adj_MAF", "ExAC_AFR_MAF", "ExAC_AMR_MAF", "ExAC_EAS_MAF", "ExAC_FIN_MAF",
         "ExAC_NFE_MAF", "ExAC_OTH_MAF", "ExAC_SAS_MAF", "CLIN_SIG", "SOMATIC", "PHENO",
         "MOTIF_NAME", "MOTIF_POS", "HIGH_INF_POS", "MOTIF_SCORE_CHANGE", "LoF",
         "LoF_filter", "LoF_flags", "LoF_info")
```

# VI. Gencode
Download the GTF file Release 42 (GRCh38.p13) : Comprehensive gene annotation	CHR	 
It contains the comprehensive gene annotation on the reference chromosomes only  
https://www.gencodegenes.org/pages/data_format.html
```{r}
gencode <-
  read.delim(#"~/Downloads/gencode.v42.annotation.gtf.gz", 
             paste0(path, "/gnomAD_skweness/other_data/gencode_filtered.vcf.gz"),
              sep = " ")
```

## Merging with DNMT3A in gnomAD
```{r}
stat_measure %>% select(IDs, X.CHROM, POS, REF, ALT, gene_in_gnomad) %>% 
  left_join(., gencode,
             by = c("gene_in_gnomad" = "gene_name"))
```





